---
title: "Who tackles the most? Looking at 2010 World Cup player stats"
author: "Rohit Bansal"
date: 2026-09-24
format: html
---

## Introduction

For this post I wanted to look at some football data and see if I could
find anything interesting using R. I always assumed defenders do most of
the tackling on a team, so I decided to check if that is actually true
using real player data from the 2010 World Cup.

I'm using the `worldcup` dataset that comes with the `faraway` R package.
It has stats for 595 players from the 2010 World Cup, like their position,
minutes played, tackles, shots, passes, and saves.

Source: `worldcup` dataset from the
[`faraway`](https://cran.r-project.org/package=faraway) package.

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(tidyverse)
library(ggplot2)
library(faraway)
```

## Looking at the data

First I just want to see what the data looks like.

```{r}
#| label: look-at-data
head(worldcup)
```

There are some players who barely played any minutes, which might mess up
the results, so I decided to only keep players who played at least 10
minutes.

```{r}
#| label: filter-data
worldcup2 <- worldcup |>
  filter(Time >= 10)

nrow(worldcup2)
```

## Tackles by position

My main question was whether defenders really do tackle more than other
positions. Since players play different amounts of time, I calculated
tackles per 90 minutes so it's a fair comparison.

```{r}
#| label: tackles-per-position
worldcup2 <- worldcup2 |>
  mutate(tackles_per_90 = Tackles / Time * 90) |>
  filter(Position != "Goalkeeper")

worldcup2 |>
  group_by(Position) |>
  summarise(avg_tackles = round(mean(tackles_per_90), 2))
```

This table already shows defenders have the highest average tackles per 90
minutes, which is what I expected. I also made a plot to see this a bit
more clearly.

```{r}
#| label: fig-tackle-boxplot
#| fig-cap: "Tackles per 90 minutes for each position (goalkeepers not included)."
worldcup2 |>
  ggplot(aes(x = Position, y = tackles_per_90)) +
  geom_boxplot() +
  labs(y = "Tackles per 90 minutes")
```

Defenders clearly tackle the most on average, but there is also a lot of
overlap between positions — some midfielders tackle just as much as some
defenders.

## Shots vs passes

I was also curious if there's a relationship between how many passes a
player makes and how many shots they take, since I'd guess attackers pass
less but shoot more.

```{r}
#| label: fig-shots-passes
#| fig-cap: "Shots vs passes for each player, coloured by position."
worldcup2 |>
  filter(Position != "Goalkeeper") |>
  ggplot(aes(x = Passes, y = Shots, color = Position)) +
  geom_point() +
  labs(x = "Passes", y = "Shots")
```

This roughly matches what I expected — a lot of the points with more shots
and fewer passes tend to be forwards, while defenders are usually on the
other side with more passes and almost no shots.

## Conclusion

Using the `worldcup` dataset, I found that defenders do tackle the most on
average compared to other positions, which confirms what I assumed at the
start. The passes vs shots plot also lined up with what I expected —
forwards shoot more and pass less, while defenders do the opposite. This
was a good first practice at loading a dataset, cleaning it a bit, and
making some plots in R.